
## R Markdown

R Markdown lets us combine narrative text, code, and output in one reproducible
document. In software engineering, this is very useful for:

- defect analysis reports
- effort estimation summaries
- model evaluation dashboards
- experiment logs for reproducibility

## Basic Structure

A typical R Markdown / Quarto document contains:

- headings and text written in Markdown
- code chunks (R, Python, etc.)
- generated output (tables, plots, metrics)

The rendering process executes the chunks and creates HTML/PDF/Word output.

## Code Chunks

Chunks start and end with triple backticks and a chunk header.

```{r}
# Example software metrics dataset
se_metrics <- data.frame(
	module = c("Auth", "Billing", "Search", "UI", "Export"),
	loc = c(220, 540, 430, 310, 270),
	cyclomatic = c(12, 25, 19, 14, 11),
	defects = c(0, 3, 2, 1, 0)
)

summary(se_metrics)
```

## Useful Chunk Options

- `echo=FALSE`: hide code, show results
- `message=FALSE`: hide package startup messages
- `warning=FALSE`: hide warnings
- `eval=FALSE`: show code but do not run it

```{r, echo=FALSE, message=FALSE, warning=FALSE}
aggregate(defects ~ module, data = se_metrics, FUN = mean)
```

## Inline Code

You can insert computed values directly in text.

Example: the total number of defects in this sample is `r sum(se_metrics$defects)`.

## Plot Example (Software Engineering)

```{r, fig.width=7, fig.height=4}
plot(
	se_metrics$loc,
	se_metrics$defects,
	xlab = "Lines of Code (LOC)",
	ylab = "Number of Defects",
	main = "Defects vs LOC by Module",
	pch = 19,
	col = "steelblue"
)
text(se_metrics$loc, se_metrics$defects, labels = se_metrics$module, pos = 3, cex = 0.8)
```

## Reproducibility Tips

- set a random seed for experiments (`set.seed(...)`)
- keep data-loading code in the document
- report metrics and plots in the same file
- avoid manual copy/paste of results into reports

## Further Information

- [Markdown Cheatsheet](https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet)
- [R Markdown by Posit](https://rmarkdown.rstudio.com/)
- [Quarto Authoring Basics](https://quarto.org/docs/authoring/)


