---
title: "tidyverse"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Tidyverse approach to Data Analysis

<https://www.tidyverse.org/>

`tidyverse` is a family of packages promoted by RStudio with the purpose of standardizing the processes in data science

## Packages included

-   `readr`
-   `tibble`
-   `dplyr` for data manipulation
-   `tidyr` for "tidy"ing data
-   `forcats`
-   `ggplot2` for plotting data
-   `purrr`
-   `stringr`
-   and the Pipe operator `%>%`

### The Pipe operator in packages magrittr and dplyr
The pipe %>% operator was first introduced in the magrittr package. Now, it is an essential part of the tidyverse library. 
It allows to concatenate a sequence of methods or calls to functions.
The operator is read from "left to right". The %>% takes the data that is on the left hand side and passes it to the function that is on the right hand side.

It comes in handy when applying a sequence of functions to data frames. 

```{r, echo=FALSE}
library(magrittr)
print("Compute the square root of the elements of a vector")
x <- 0:5
x
sqrt(x)

"  Using the pipe %>% ..." %>% print()
x %>% sqrt()

x %>% summary()

library(dplyr)
mtcars %>% select(mpg, cyl, hp) %>% filter(cyl > 4) %>% summary()

mtcars_2 <- mtcars %>% select(mpg, cyl, hp) %>% filter(cyl > 4) %>% head()
mtcars_2


```

### Criticisms to this approach

-   <https://towardsdatascience.com/a-thousand-gadgets-my-thoughts-on-the-r-tidyverse-2441d8504433>
-   <https://github.com/matloff/TidyverseSkeptic/blob/master/READMEFull.md>
-   <https://www.r-bloggers.com/2019/12/why-i-dont-use-the-tidyverse/>
-   Jared P. Lander, R for Everyone. Advanced Analytics and Graphics, 2nd ed., Addison-Wesley, 2017
