
# Advanced Preprocessing Techniques

This chapter complements basic preprocessing with techniques that are often
required in realistic software engineering datasets.

## Feature Selection

This technique consists of selecting the most relevant attributes. The need for applying FS includes the following points:

  + A reduced volume of data allows different data mining or searching techniques to be applied.

  + Irrelevant and redundant attributes can generate less accurate and more complex models. Furthermore, data mining algorithms can be executed faster.

  + It is possible to avoid the collection of data for those irrelevant and redundant attributes in the future.



FS algorithms designed with different evaluation criteria broadly fall into two categories:

  + The *filter model* relies on general characteristics of the data to evaluate and select feature subsets without involving any data mining algorithm.

  + The *wrapper model* requires one predetermined mining algorithm and uses its performance as the evaluation criterion. It searches for features better suited to the mining algorithm aiming to improve mining performance, but it also tends to be more computationally expensive than filter model [11, 12].



## Instance Selection

Instance selection aims to reduce dataset size while preserving predictive
performance. In software engineering datasets, this is useful when duplicated
or near-duplicated modules inflate training time and bias model behavior.

Typical strategies include:

- distance-based prototype selection for nearest-neighbor methods
- noise filtering before model training
- uncertainty-based sampling in iterative workflows

In practice, evaluate instance selection with repeated resampling and compare
both performance and model stability against a no-selection baseline.


## Missing Data Imputation


Missing data imputation should be treated as part of the training pipeline, not
as a one-off preprocessing script. This avoids data leakage and improves
reproducibility.

Recommended workflow:

- analyze missingness patterns first (MCAR, MAR, MNAR assumptions)
- choose imputation strategy by variable type and modeling objective
- fit imputation using only training folds
- validate impact on downstream metrics

Common options in modern R workflows include median/mode imputation through
`recipes`, nearest-neighbor imputation, and model-based methods for richer
datasets.


## Encoding and transformation pipelines

Advanced workflows usually include mixed variable types and skewed numeric
distributions. A practical recipe often combines:

- dummy encoding for categorical variables
- variance filtering for unstable predictors
- normalization for distance/regularized models
- optional log/Yeo-Johnson transforms for skewed features

```{r message=FALSE, warning=FALSE}
library(tidymodels)

df <- read.csv("./datasets/defectPred/unified/Unified-file.csv", stringsAsFactors = FALSE)
df <- df[, c("McCC", "CLOC", "PDA", "PUA", "LLOC", "LOC", "bug")]
df$Defective <- factor(ifelse(df$bug > 0, "Y", "N"))
df$bug <- NULL

rec_adv <- recipe(Defective ~ ., data = df) |>
  step_zv(all_predictors()) |>
  step_nzv(all_predictors()) |>
  step_YeoJohnson(all_numeric_predictors()) |>
  step_normalize(all_numeric_predictors())

rec_adv |> prep() |> juice() |> dplyr::glimpse()
```


## Temporal splits and concept drift

For SE prediction tasks, data distributions can change across releases
(concept drift). Prefer temporal evaluation when date/order information is
available.

```{r message=FALSE, warning=FALSE}
set.seed(10)
n <- nrow(df)
cutoff <- floor(0.7 * n)

train_time <- df[1:cutoff, ]
test_time <- df[(cutoff + 1):n, ]

summary(train_time$LOC)
summary(test_time$LOC)
```


## Deployment consistency

A frequent production issue is mismatch between training-time preprocessing and
inference-time preprocessing. Persist preprocessing objects (recipes, lookup
tables, selected features) and reuse them unchanged in deployment.
