
# Regression {#regression}

## Learning Objectives and Evaluation Lens

- **Objective**: model and interpret numeric software engineering outcomes (for example, effort).
- **Data context**: project/module predictors with a continuous response variable.
- **Validation**: train/test split (prefer temporal split when available) and diagnostic checks.
- **Primary metrics**: MAE, RMSE, and $R^2$.
- **Common pitfalls**: extrapolation beyond observed ranges, nonlinearity, heteroscedasticity, and leakage in preprocessing.

## Validation strategy for SE regression

For effort/time prediction, random splits are often optimistic when records are
project-correlated or time-ordered.

- prefer temporal splits (train on older projects/releases)
- use group-aware splits when many rows come from the same project
- avoid tuning and reporting on the same folds

```{r eval=FALSE}
# Temporal split sketch
# telecom1 <- telecom1 |> dplyr::arrange(project_end_date)
# split_idx <- floor(0.8 * nrow(telecom1))
# train <- telecom1[1:split_idx, ]
# test  <- telecom1[(split_idx + 1):nrow(telecom1), ]
```

### Nested resampling (regression)

Nested resampling gives less biased performance estimates for tuned models.

```{r eval=FALSE}
# outer <- rsample::vfold_cv(train, v = 5)
# for each outer split:
#   inner <- rsample::vfold_cv(training(outer_split), v = 5)
#   tune model on inner folds
#   refit best model on training(outer_split)
#   score on testing(outer_split)
```

## Linear Regression modeling

  - _Linear Regression_ is one of the oldest and most known predictive methods. As its name says, the idea is to try to fit a linear equation between a dependent variable and an independent, or explanatory, variable. The idea is that the independent variable $x$ is something the experimenter controls and the dependent variable $y$ is something that the experimenter measures. The line is used to predict the value of $y$ for a known value of $x$. The variable $x$ is the predictor variable and $y$ the response variable.
  
  - _Multiple linear regression_ uses 2 or more independent variables for building a model. See  <https://www.wikipedia.org/wiki/Linear_regression>.
  
  - First proposed many years ago but still very useful for software engineering tasks such as effort estimation and defect-count modeling.
  
  - The equation takes the form $\hat{y}=b_0+b_1 * x$
  - The method used to choose the values $b_0$ and $b_1$ is to minimize the sum of the squares of the residual errors.

### Regression: Software Effort Example

The example below uses the `Telecom1` software effort dataset.

```{r warning=FALSE, message=FALSE}
telecom1 <- read.table("./datasets/effortEstimation/Telecom1.csv", sep=",", header=TRUE, stringsAsFactors=FALSE, dec = ".")

par(mfrow=c(1,2))
hist(telecom1$size, col="blue", breaks=12, main="Project size")
hist(telecom1$effort, col="blue", breaks=12, main="Project effort")

plot(telecom1$size, telecom1$effort, pch=1, col="blue", cex=0.8,
  xlab="Size", ylab="Effort")
lm1 <- lm(effort ~ size, data = telecom1)
abline(lm1, col="red", lwd=3)

plot(telecom1$size, lm1$residuals, col="blue", pch=1, cex=0.8,
  xlab="Size", ylab="Residuals")
abline(c(0,0), col="red", lwd=2)
qqnorm(lm1$residuals)
qqline(lm1$residuals)
```

### Simple Linear Regression

- Given two variables $Y$ (response) and $X$ (predictor), the assumption is that there is an approximate ($\approx$) *linear* relation between those variables. 
- The mathematical model of the observed data is described as (for the case of simple linear regression):
$$ Y \approx \beta_0 + \beta_1 X$$

- the parameter $\beta_0$ is named the *intercept* and $\beta_1$ is the  *slope*
- Each observation can be modeled as

$$y_i = \beta_0 + \beta_1 x_i + \epsilon_i;
\epsilon_i \sim N(0,\sigma^2)$$ 
- $\epsilon_i$ is the *error*
- This means that the variable $y$ is normally distributed:
$$ y_i \sim N( \beta_0 + \beta_1 x_i, \sigma^2) $$


- The *predictions* or *estimations* of this model are obtained by a linear equation of the form $\hat{Y}=\hat{\beta_0} + \hat{\beta}_1X$, that is, each new prediction is computed with
$$\hat{y}_i = \hat{\beta}_0 + \hat{\beta}_1x_i $$.
- The actual parameters $\beta_0$ and $\beta_1$ are unknown
- The parameters $\hat{\beta}_0$ and $\hat{\beta}_1$ of the linear equation can be estimated with different methods. 

### Least Squares

- One of the most used methods for computing $\hat{\beta}_0$ and $\hat{\beta}_1$ is the criterion of "least squares" minimization. 
- The data is composed of $n$ pairs of observations $(x_i, y_i)$
- Given an observation $y_i$ and its corresponding estimation $\hat{y_i})$ the *residual* $e_i$ is defined as $$e_i= y_i - \hat{y_i}$$
- the Residual Sum of Squares is defined as $$RSS=e_1^2+\dots + e_i^2+\dots+e_n^2$$
- the Least Squares Approach minimizes the RSS
- as a result of that minimization, it can be obtained, by means of calculus, the estimation of $\hat{\beta}_0$ and $\hat{\beta}_1$ as $$\hat{\beta}_1=\frac{\sum_{i=1}^{n}{(x_i-\bar{x})(y_i-\bar{y})}}{\sum_{i=1}^{n}(x_i-\bar{x})^2}$$ and $$\hat{\beta}_0=\bar{y}-\hat{\beta}_1\bar{x} $$ where $\bar{y}$ and $\bar{x}$ are the sample means.
- the variance $\sigma^2$ is estimated by 
$$\hat\sigma^2 = {RSS}/{(n-2)}$$ where n is the number of observations 
- The *Residual Standard Error* is defined as $$RSE = \sqrt{{RSS}/{(n-2)}}$$
- The equation $$ Y = \beta_0 + \beta_1 X + \epsilon$$ defines the linear model, i.e., the *population regression line*
- The *least squares line* is $\hat{Y}=\hat{\beta_0} + \hat{\beta}_1X$
- *Confidence intervals* are computed using the *standard errors* of the intercept and the slope.
- The $95\%$ confidence interval for the slope is computed as $$[\hat{\beta}_1 - 2 \cdot SE(\hat{\beta}_1), \hat{\beta}_1+SE(\hat{\beta}_1)]$$ 
- where $$ SE(\hat{\beta}_1) = \sqrt{\frac{\sigma^2}{\sum_{i=1}^{n}(x_i-\bar{x})^2}}$$

### Linear regression in R

The following are the basic commands in R:

- The basic function is `lm()`, that returns an object with the model. 
- Other commands: `summary` prints out information about the regression, `coef` gives the coefficients for the linear model, `fitted` gives the predictd value of $y$ for each value of $x$, `residuals` contains the differences between observed and fitted values. 
- [`predict`](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/predict.lm.html) will generate predicted values of the response for the values of the explanatory variable. 
  
  
## Linear Regression Diagnostics

  - Several plots help to evaluate the suitability of the linear regression 
    + *Residuals vs fitted*: The residuals should be randomly distributed around the horizontal line representing a residual error of zero; that is, there should not be a distinct trend in the distribution of points. 
    + *Standard Q-Q plot*: residual errors are normally distributed
    + *Square root of the standardized residuals vs the fitted values*: there should be no obvious trend. This plot is similar to the residuals versus fitted values plot, but it uses the square root of the standardized residuals. 
    + *Leverage*: measures the importance of each point in determining the regression result. Smaller values means that removing the observation has little effect on the regression result. 


### Simulation example

#### Simulate a dataset
```{r}
set.seed(3456)
# equation is  y = -6.6 + 0.13 x +e
# range x 100,400
a <- -6.6
b <- 0.13
num_obs <- 60
xmin <- 100
xmax <- 400
x <- sample(seq(from=xmin, to = xmax, by =1), size= num_obs, replace=FALSE)

sderror <- 9 # sigma for the error term in the model
e <- rnorm(num_obs, 0, sderror) 

y <- a + b * x + e


newlm <- lm(y~x)
summary(newlm)

cfa1 <- coef(newlm)[1]
cfb2 <- coef(newlm)[2]
plot(x,y, xlab="x axis", ylab= "y axis", xlim = c(xmin, xmax), ylim = c(0,60), sub = "Line in black is the actual model")
title(main = paste("Line in blue is the Regression Line for ", num_obs, " points."))

abline(a = cfa1, b = cfb2, col= "blue", lwd=3)
abline(a = a, b = b, col= "black", lwd=1) #original line

```


##### Subset a set of points from the same sample
```{r}

# sample from  the same  x     to compare least squares lines 
# change the denominator in newsample to see how the least square lines changes accordingly. 
newsample <- as.integer(num_obs/8) # number of pairs x,y

idxs_x1 <- sample(1:num_obs, size = newsample, replace = FALSE) #sample indexes
x1 <- x[idxs_x1]
e1 <- e[idxs_x1]
y1 <- a + b * x1 + e1
xy_obs <- data.frame(x1, y1)
names(xy_obs) <- c("x_obs", "y_obs")

newlm1 <- lm(y1~x1)
summary(newlm1)

cfa21 <- coef(newlm1)[1]
cfb22 <- coef(newlm1)[2]

plot(x1,y1, xlab="x axis", ylab= "y axis", xlim = c(xmin, xmax), ylim = c(0,60))
title(main = paste("New line in red with ", newsample, " points in sample"))

abline(a = a, b = b, col= "black", lwd=1)  # True line
abline(a = cfa1, b = cfb2, col= "blue", lwd=1)  #sample
abline(a = cfa21, b = cfb22, col= "red", lwd=2) #new line
```


##### Compute a confidence interval on the original sample regression line

```{r}

newx <- seq(xmin, xmax)
ypredicted <- predict(newlm, newdata=data.frame(x=newx), interval= "confidence", level= 0.90, se = TRUE)

plot(x,y, xlab="x axis", ylab= "y axis", xlim = c(xmin, xmax), ylim = c(0,60))
# points(x1, fitted(newlm1))
abline(newlm)

lines(newx,ypredicted$fit[,2],col="red",lty=2)
lines(newx,ypredicted$fit[,3],col="red",lty=2)

# Plot the residuals or errors
ypredicted_x <- predict(newlm, newdata=data.frame(x=x))
plot(x,y, xlab="x axis", ylab= "y axis", xlim = c(xmin, xmax), ylim = c(0,60), sub = "", pch=19, cex=0.75)
title(main = paste("Residuals or errors", num_obs, " points."))
abline(newlm)
segments(x, y, x, ypredicted_x)

```

##### Take another sample from the model and explore
```{r}
# equation is  y = -6.6 + 0.13 x +e
# range x 100,400
num_obs <- 35
xmin <- 100
xmax <- 400
x3 <- sample(seq(from=xmin, to = xmax, by =1), size= num_obs, replace=FALSE)
sderror <- 14 # sigma for the error term in the model
e3 <- rnorm(num_obs, 0, sderror) 

y3 <- a + b * x3 + e3

newlm3 <- lm(y3~x3)
summary(newlm3)

cfa31 <- coef(newlm3)[1]
cfb32 <- coef(newlm3)[2]
plot(x3,y3, xlab="x axis", ylab= "y axis", xlim = c(xmin, xmax), ylim = c(0,60))
title(main = paste("Line in red is the Regression Line for ", num_obs, " points."))
abline(a = cfa31, b = cfb32, col= "red", lwd=3)
abline(a = a, b = b, col= "black", lwd=2) #original line
abline(a = cfa1, b = cfb2, col= "blue", lwd=1) #first sample

# confidence intervals for the new sample

newx <- seq(xmin, xmax)
ypredicted <- predict(newlm3, newdata=data.frame(x3=newx), interval= "confidence", level= 0.90, se = TRUE)

lines(newx,ypredicted$fit[,2],col="red",lty=2, lwd=2)
lines(newx,ypredicted$fit[,3],col="red",lty=2, lwd=2)

```


### Diagnostics fro assessing the regression line

#### Residual Standard Error
- It gives us an idea of the typical or average error of the model. It is the estimated standard deviation of the residuals.

#### $R^2$ statistic
- This is the proportion of variability in the data that is explained by the model. Best values are those close to 1.


## Multiple Linear Regression

### Partial Least Squares
- If several predictors are highly correlated, the least squares approach has high variability. 
- PLS finds linear combinations of the predictors, that are called *components* or *latent* variables. 



## Linear regression in Software Effort estimation

Fitting a linear model to log-log
  - the predictive power equation is $y= e^{b_0}*x^{b_1}$, ignoring the bias corrections. Note: depending how the error term behaves we could try another general linear model (GLM) or other model that does not rely on the normality of the residuals (quantile regression, etc.)
  - First, we are fitting the model to the whole dataset. But it is not the right way to do it, because of overfitting.


```{r warning=FALSE, message=FALSE}
library(foreign)
china <- read.arff("./datasets/effortEstimation/china.arff")
china_size <- china$AFP
summary(china_size)
china_effort <- china$Effort
summary(china_effort)
par(mfrow=c(1,2))
hist(china_size, col="blue", xlab="Adjusted Function Points", main="Distribution of AFP")
hist(china_effort, col="blue",xlab="Effort", main="Distribution of Effort")
boxplot(china_size)
boxplot(china_effort)
qqnorm(china_size)
qqline(china_size)
qqnorm(china_effort)
qqline(china_effort)
```
  
Applying the `log` function (it computes natural logarithms, base $e$)

```{r, echo=FALSE}
par(mfrow=c(1,2))
logchina_size = log(china_size)
hist(logchina_size, col="blue", xlab="log Adjusted Function Points", main="Distribution of log AFP")
logchina_effort = log(china_effort)
hist(logchina_effort, col="blue",xlab="Effort", main="Distribution of log Effort")
qqnorm(logchina_size)
qqnorm(logchina_effort)
```
  
  
  
```{r}
linmodel_logchina <- lm(logchina_effort ~ logchina_size)
par(mfrow=c(1,1))
plot(logchina_size, logchina_effort)
abline(linmodel_logchina, lwd=3, col=3)
par(mfrow=c(1,2))
plot(linmodel_logchina, ask = FALSE)
linmodel_logchina
```



## Exercise: Build a Regression Model for Effort Estimation {#sec-exercise-regression}

This exercise asks you to build a complete linear regression pipeline — from loading a dataset to evaluating predictions on held-out data — and document the process in a reproducible R Markdown / Quarto document.

### Part A — Required

1. **Create a new project and document.** Open RStudio (or Positron/VS Code) and start a new `.qmd` or `.Rmd` file.

2. **Choose a dataset.** Go to `datasets/effortEstimation/` and pick one ARFF or CSV file (e.g., `boehm.arff`, `albrecht.arff`, `miyazaki94.arff`). The file must contain at least a `size` column and an `effort` column.

3. **Load the data.**
```{r ex-load, eval=FALSE}
library(foreign)
path <- file.path(getwd(), "datasets/effortEstimation/albrecht.arff")
df   <- read.arff(path)
```

4. **Explore the data.** Report summary statistics and visualise the `size`–`effort` relationship. Check for outliers and missing values.

5. **Assess normality.** Plot histograms and Q-Q plots. Use the Shapiro-Wilk test if needed. Transform variables (e.g., `log`) if the distributions are heavily skewed.

6. **Check correlation.** Compute the Pearson / Spearman correlation between `size` (or `log(size)`) and `effort` (or `log(effort)`).

7. **Split into train and test sets.**
```{r ex-split, eval=FALSE}
set.seed(42)
n      <- nrow(df)
idx    <- sample(seq_len(n), size = floor(0.8 * n))
train  <- df[ idx, ]
test   <- df[-idx, ]
write.csv(train, "train.csv", row.names = FALSE)
write.csv(test,  "test.csv",  row.names = FALSE)
```

8. **Build the linear model** on the training data and plot the fitted line.

9. **Evaluate on the test set.** Compute MAE, MdAE, and Pred(25) on back-transformed predictions (see @sec-eval-regression). Avoid MMRE as the primary metric — it is biased toward underestimating models because overestimates and underestimates of the same magnitude receive different penalties [@Shepperd2012]. MdMRE is acceptable as a supplementary measure.

10. **Write conclusions.** Discuss model fit, residuals, and whether the assumptions of linear regression hold for this dataset.

### Part B — Optional Extensions

- Apply an additional model (e.g., random forest via `ranger`, or regularised regression via `glmnet`) and compare evaluation metrics.
- Apply a stratified data split to ensure both splits cover the full range of project sizes.
- Report confidence intervals for key coefficients.
- Wrap the data and analysis in a small R package (see @sec-rpackages).

## Further Information

- [The New Statistics with R (Andy Hector)](https://global.oup.com/academic/product/the-new-statistics-with-r-9780198729051)
- [An Introduction to R (R Core Team)](https://cran.r-project.org/doc/manuals/r-release/R-intro.html)
- [Practical Data Science with R (Nina Zumel, John Mount)](https://www.manning.com/books/practical-data-science-with-r)
- [An Introduction to Statistical Learning with Applications in R (ISLR)](https://www.statlearning.com/)
