# Feature Selection Example

This chapter presents a modern feature-selection workflow using `tidymodels`.
The goal is to keep the process explicit and reproducible:

- split data into training and test subsets
- define feature filters in a `recipe`
- fit a model through a `workflow`
- inspect variable importance

```{r setup_fs, message=FALSE, warning=FALSE}
library(tidymodels)
library(vip)
library(foreign)

set.seed(10)
kc1 <- read.arff("./datasets/defectPred/D1/KC1.arff")

split <- initial_split(kc1, prop = 0.75, strata = Defective)
training <- training(split)
testing <- testing(split)
```

## Feature filtering with recipes

`recipes` supports common feature-selection and cleaning steps:

- `step_zv()` removes zero-variance predictors
- `step_nzv()` removes near-zero-variance predictors
- `step_corr()` removes highly correlated numeric predictors

```{r fs_recipe, message=FALSE, warning=FALSE}
rec <- recipe(Defective ~ ., data = training) |>
  step_zv(all_predictors()) |>
  step_nzv(all_predictors()) |>
  step_corr(all_numeric_predictors(), threshold = 0.90)

rec_prep <- prep(rec)
juice(rec_prep) |> dplyr::glimpse()
```

## Train a model after feature filtering

```{r fs_model, message=FALSE, warning=FALSE}
rf_spec <- rand_forest(trees = 500, min_n = 5) |>
  set_mode("classification") |>
  set_engine("ranger", importance = "impurity")

wf <- workflow() |>
  add_recipe(rec) |>
  add_model(rf_spec)

rf_fit <- fit(wf, data = training)
rf_fit
```

## Evaluate on the test set

```{r fs_eval, message=FALSE, warning=FALSE}
pred_cls <- predict(rf_fit, testing, type = "class")
pred_prb <- predict(rf_fit, testing, type = "prob")

eval_tbl <- bind_cols(testing, pred_cls, pred_prb)

metrics(eval_tbl, truth = Defective, estimate = .pred_class)
conf_mat(eval_tbl, truth = Defective, estimate = .pred_class)
```

## Variable importance

```{r fs_importance, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
rf_fit |>
  extract_fit_parsnip() |>
  vip(num_features = 15)
```
