# Evaluation of Unsupervised Models

## Learning Objectives and Evaluation Lens

- **Objective**: evaluate whether unsupervised structure is both statistically coherent and practically useful.
- **Primary metrics**: silhouette, CH index, purity, ARI.
- **Interpretation rule**: internal quality and external usefulness should both be considered.
- **Common pitfalls**: selecting $k$ once without sensitivity analysis and over-interpreting unstable clusters.

In unsupervised learning we do not train against target labels, so evaluation
must focus on cluster structure quality and business usefulness.

In software engineering, a common use case is grouping modules/components by
their static metrics (size, complexity, coupling) to identify *risk profiles*
for testing and quality assurance.

## Example: Clustering Software Modules (UnifiedBugDataset)

We use `Unified-file.csv` from UnifiedBugDataset 1.2 (2019), which merges
multiple defect datasets (PROMISE, Eclipse, Bug Prediction Dataset,
Bugcatchers, GitHub bug datasets). Compared to using only NASA subsets (e.g.,
KC1/JM1), this is a broader and more up-to-date benchmark.

Even though clustering is unsupervised, the dataset includes a `bug` field that
we use *after* clustering as an external validation signal.

```{r setup_unsup_eval, message=FALSE, warning=FALSE, include=FALSE}
has_cluster <- requireNamespace("cluster", quietly = TRUE)

if (!has_cluster) {
	message("Package 'cluster' is not installed; silhouette metric will be skipped.")
}

set.seed(2026)
```

```{r load_data_unsup_eval, message=FALSE, warning=FALSE}
unified <- read.csv("./datasets/defectPred/unified/Unified-file.csv", stringsAsFactors = FALSE)

# External validation label (binary): defective if bug count > 0.
truth <- factor(ifelse(unified$bug > 0, "buggy", "clean"), levels = c("clean", "buggy"))

# Use numeric software metrics only, excluding identifiers and bug label.
drop_cols <- c("ID", "Name", "LongName", "Parent", "bug")
X <- unified[, setdiff(names(unified), drop_cols)]
X <- X[, sapply(X, is.numeric), drop = FALSE]

# Remove incomplete rows and keep labels aligned.
ok <- complete.cases(X)
X <- X[ok, , drop = FALSE]
truth <- truth[ok]

# Remove degenerate columns that break scaling/kmeans.
var_ok <- vapply(X, function(col) {
	v <- var(col, na.rm = TRUE)
	is.finite(v) && v > 0
}, logical(1))
X <- X[, var_ok, drop = FALSE]

# Silhouette requires pairwise distances (O(n^2)); cap sample size for speed.
max_n <- 2500
if (nrow(X) > max_n) {
	idx <- sample(seq_len(nrow(X)), size = max_n)
	X <- X[idx, , drop = FALSE]
	truth <- truth[idx]
}

# Scale metrics to avoid size-dominated clusters.
X_scaled <- scale(X)
X_scaled <- X_scaled[, colSums(is.finite(X_scaled)) == nrow(X_scaled), drop = FALSE]

dim(X_scaled)
```

## Internal Evaluation Metrics

Internal metrics evaluate cluster compactness and separation without labels.

- Total within-cluster sum of squares (`tot.withinss`): lower is better
- Average silhouette width: higher is better
- Calinski-Harabasz index: higher is better

```{r internal_metrics_functions}
calinski_harabasz <- function(x, clusters) {
	x <- as.matrix(x)
	clusters <- as.factor(clusters)
	n <- nrow(x)
	k <- nlevels(clusters)

	overall_center <- colMeans(x)

	# Within-cluster sum of squares
	wss <- 0
	# Between-cluster sum of squares
	bss <- 0

	for (cl in levels(clusters)) {
		idx <- which(clusters == cl)
		xk <- x[idx, , drop = FALSE]
		nk <- nrow(xk)
		ck <- colMeans(xk)

		wss <- wss + sum((xk - matrix(ck, nrow = nk, ncol = ncol(xk), byrow = TRUE))^2)
		bss <- bss + nk * sum((ck - overall_center)^2)
	}

	(bss / (k - 1)) / (wss / (n - k))
}

avg_silhouette <- function(x, clusters) {
	if (!has_cluster) return(NA_real_)
	sil <- cluster::silhouette(as.integer(as.factor(clusters)), dist(x))
	mean(sil[, "sil_width"])
}
```

```{r compare_k_values, message=FALSE, warning=FALSE}
ks <- 2:6

internal_tbl <- lapply(ks, function(k) {
	km <- kmeans(X_scaled, centers = k, nstart = 25)
	data.frame(
		k = k,
		total_withinss = km$tot.withinss,
		avg_silhouette = avg_silhouette(X_scaled, km$cluster),
		calinski_harabasz = calinski_harabasz(X_scaled, km$cluster)
	)
})

internal_tbl <- do.call(rbind, internal_tbl)
internal_tbl
```

## External Evaluation (Using Defect Labels Only for Assessment)

External metrics compare clusters with known classes. This does not make the
training supervised; it only checks whether clusters align with practical
categories of interest.

- Purity: proportion of modules assigned to the majority class in each cluster
- Adjusted Rand Index (ARI): agreement corrected for chance

```{r external_metrics_functions}
purity_score <- function(truth, pred) {
	tab <- table(truth, pred)
	sum(apply(tab, 2, max)) / sum(tab)
}

adjusted_rand_index <- function(truth, pred) {
	tab <- table(truth, pred)
	n <- sum(tab)

	a <- rowSums(tab)
	b <- colSums(tab)

	sum_choose <- sum(choose(tab, 2))
	expected <- (sum(choose(a, 2)) * sum(choose(b, 2))) / choose(n, 2)
	max_index <- 0.5 * (sum(choose(a, 2)) + sum(choose(b, 2)))

	denom <- max_index - expected
	if (denom == 0) return(0)

	(sum_choose - expected) / denom
}
```

```{r external_metrics_example, message=FALSE, warning=FALSE}
# Choose k by best silhouette when available, otherwise k=2 as baseline.
if (all(is.na(internal_tbl$avg_silhouette))) {
	k_best <- 2
} else {
	k_best <- internal_tbl$k[which.max(internal_tbl$avg_silhouette)]
}

km_best <- kmeans(X_scaled, centers = k_best, nstart = 25)

external_tbl <- data.frame(
	k = k_best,
	purity = purity_score(truth, km_best$cluster),
	adjusted_rand_index = adjusted_rand_index(truth, km_best$cluster)
)

external_tbl
```

## SE Interpretation of the Result

In this context, useful unsupervised evaluation means:

- Clusters are internally coherent (higher silhouette / CH, lower WSS)
- Clusters have interpretable engineering meaning
	- for example, one cluster might contain high complexity and size metrics,
		indicating modules requiring stronger testing effort
- If external validation is available, non-trivial purity/ARI suggests that
	structural metric groups are related to defect proneness

This evaluation approach supports practical decisions such as risk-based test
prioritization, review allocation, and focused refactoring.

## Important topics often missing

- cluster stability under resampling (same data, different seeds)
- sensitivity to feature scaling choices
- sensitivity to parameter choices ($k$, `eps`, `MinPts`)
- actionable interpretation quality (can teams act on cluster descriptions?)
