
# Further Classification Models 

## Learning Objectives and Evaluation Lens

- **Objective**: handle classification settings that go beyond single-label fully supervised learning.
- **Data context**: multilabel corpora and partially labeled datasets.
- **Validation**: strict separation of labeled/unlabeled/test data with leakage controls.
- **Primary metrics**: macro/micro F1 for multilabel tasks; confusion-based metrics for SSL classification.
- **Common pitfalls**: pseudo-label confirmation bias, class imbalance, and distribution mismatch.


## Multilabel classification

Some datasets, for example reviews of applications and mobile application
repositories such as the App Store or Google Play, contain reviews that can
have several labels at the same time (e.g., bugs, feature requests, UI issues).
Similarly, a software issue may belong to multiple components or concern
multiple quality attributes simultaneously.

In **multilabel classification**, each instance can be assigned to more than
one class. Common approaches are:

- **Binary Relevance (BR)**: train one independent binary classifier per label.
  Simple but ignores label co-occurrence.
- **Classifier Chains (CC)**: extend BR by passing previous predictions as
  features to the next classifier, capturing label dependencies.
- **Label Powerset (LP)**: treat each unique label combination as a single
  class; exact but combinatorially expensive.

In R, the `mldr` package provides data structures and learners for multilabel
classification. For a tidy workflow, `mlr3` with its multilabel task type is an
alternative. Evaluation uses **Hamming loss** (fraction of wrong label
assignments), **macro/micro F1**, and **subset accuracy** (all labels correct).

```{r eval=FALSE}
# Sketch with mldr
# library(mldr)
# mldr_dataset <- mldr_from_dataframe(df, labelIndices = label_col_indices)
# summary(mldr_dataset)
```


## Ordinal Classification

Many SE outcomes are not nominal but **ordinal**: defect severity
(low < medium < high < critical), code review priority, or technical debt
level. Standard classifiers treat classes as exchangeable and ignore the
ordering, which can inflate nominal error rates on practically small mistakes.

Approaches:
- **Ordinal regression** (proportional-odds model): extends logistic regression
  with monotonicity constraints across thresholds.
- **Threshold decomposition**: convert the ordinal problem into a series of
  binary problems using cumulative link models.
- **Cost-sensitive learning**: assign asymmetric misclassification costs that
  reflect the ordinal distance.

In R, the `ordinal` package implements cumulative link (mixed) models.

```{r eval=FALSE}
# library(ordinal)
# m <- clm(severity ~ loc + complexity, data = train_df)
# predict(m, newdata = test_df)
```


## Weak and Distant Supervision

In SE datasets, labels are often derived from heuristics rather than manual
annotation — for example, a file is labelled "defective" if a commit message
contains the word *fix* or *bug* (SZZ approach), or a commit is labelled as
a "feature addition" based on keyword matching. These automatically generated,
imperfect labels are called **weak labels** or **distant labels**.

**Consequences:**
- Label noise can substantially degrade model performance.
- Models may learn to predict the labelling heuristic rather than the
  underlying concept.
- Standard evaluation metrics are optimistic if the test set uses the same
  noisy labels as training.

**Mitigations:**
- Use a small set of carefully hand-labelled instances to validate label
  quality.
- Apply label-noise-robust learning algorithms (e.g., learning with noisy
  labels via bootstrapping or noise-tolerant loss functions).
- Report sensitivity analyses showing performance variation under different
  label-cleaning thresholds.


## Active Learning

**Active learning** addresses the situation where unlabeled data is abundant
but labeling is expensive (requiring expert code review, manual bug
classification, or security audit). The learner iteratively queries an
*oracle* (typically a human expert) for the labels of the most informative
instances.

Common query strategies:
- **Uncertainty sampling**: query the instance on which the current model is
  least confident.
- **Query by committee**: train a committee of models; query the instance with
  the highest disagreement.
- **Expected model change**: query the instance that would most change the model.

In SE, active learning has been applied to defect labeling, code review
prioritization, and test-case selection. The key advantage is that fewer
expert-labeled examples are needed to reach a target accuracy.

```{r eval=FALSE}
# Active learning loop sketch (uncertainty sampling)
# 1. Start with small labeled set L, large unlabeled pool U
# 2. Train classifier on L
# 3. Compute prediction entropy on U
# 4. Select top-k most uncertain instances → send to oracle
# 5. Add newly labeled instances to L, remove from U
# 6. Repeat until budget exhausted
```


## Semi-supervised Learning

Self-train a model on semi-supervised data.

### Practical caveats

Semi-supervised learning can improve performance when unlabeled data matches
the same distribution as labeled data, but it can also amplify errors.

Key risks and guardrails:

1. **Confirmation bias**: wrong pseudo-labels can reinforce themselves.
      : Use confidence thresholds and add pseudo-labels gradually.
2. **Distribution mismatch**: unlabeled pool may differ from labeled data.
      : Monitor drift and evaluate on a clean held-out set.
3. **Overconfidence**: probabilities may be poorly calibrated.
      : Prefer calibrated models or conservative thresholds.
4. **Leakage**: test data must never be used in pseudo-label loops.
      : Keep strict train/validation/test separation.

The legacy examples in this chapter used `DMwR2::SelfTrain`, which is no longer
actively maintained. A standard modern alternative in R workflows is to combine
`tidymodels` with a pseudo-labeling loop.

```{r ssl, message=FALSE, warning=FALSE, eval = FALSE}
library(tidymodels)

## Small example with the Iris classification data set
set.seed(123)
data(iris)

## Split into train / test
idx <- sample(seq_len(nrow(iris)), 100)
tr <- iris[idx, ]
ts <- iris[-idx, ]

## Hide labels for a large subset of the training data
tr_ssl <- tr
nas <- sample(seq_len(nrow(tr_ssl)), 70)
tr_ssl$Species[nas] <- NA

## Start from labeled rows only
labeled <- tr_ssl |> dplyr::filter(!is.na(Species))
unlabeled <- tr_ssl |> dplyr::filter(is.na(Species))

## Base learner
wf <- workflow() |>
      add_recipe(recipe(Species ~ ., data = labeled)) |>
      add_model(decision_tree() |> set_engine("rpart") |> set_mode("classification"))

## Pseudo-labeling loop (single pass for teaching clarity)
fit0 <- fit(wf, data = labeled)
probs <- predict(fit0, unlabeled, type = "prob")
preds <- predict(fit0, unlabeled, type = "class")

## Keep only high-confidence pseudo-labels
conf <- apply(as.matrix(probs), 1, max)
selected <- conf >= 0.9

pseudo <- unlabeled[selected, ]
pseudo$Species <- preds$.pred_class[selected]

augmented <- dplyr::bind_rows(labeled, pseudo)
fit_ssl <- fit(wf, data = augmented)

## Evaluate on held-out test set
pred_ssl <- predict(fit_ssl, ts, type = "class")
yardstick::conf_mat_vec(ts$Species, pred_ssl$.pred_class)
```


