
# Time Series


Many sources of information are time related. For example, data from Software Configuration Management (SCM) such as Git, [GitHub](http://www.github.com)) systems or Dashboards such as [Grimoire Lab](https://chaoss.github.io/grimoirelab/) from [Bitergia](https://bitergia.com/) or [SonarQube](https://www.sonarqube.org/) 


With MetricsGrimore or SonarQube we can extract datasets or dump of databases. For example, a dashboard for the OpenStack project is located at [http://activity.openstack.org/dash/browser/](http://activity.openstack.org/dash/browser/) and provides datasets as MySQL dumps or JSON files.

With R we can read a JSON file as follows:

```{r}
library(jsonlite)
# Get the JSON data 
# gm <- fromJSON("http://activity.openstack.org/dash/browser/data/json/nova.git-scm-rep-evolutionary.json")
gm <- fromJSON('./datasets/timeSeries/nova.git-scm-rep-evolutionary.json')
str(gm)
```

Now we can use time series packages. First, after loading the libraries, we need to create a time series object.

```{r}
# TS libraries
library(xts)
library(forecast)

# Library to deal with dates
library(lubridate)

# Ceate a time series object
gmts <- xts(gm$commits,seq(ymd('2010-05-22'),ymd('2015-11-16'), by = '1 week'))

# TS Object
str(gmts)

head(gmts, 3)
```


Visualise the time series object
```{r}
plot(gmts)
```

Arima model:
```{r}
fit <- auto.arima(gmts)
fit
```

```{r}
forecast(fit, 5)
plot(forecast(fit, 5))
```


## SE-Specific Time Series Patterns

Software engineering produces several prototypical time series patterns worth
recognizing before applying generic forecasting methods:

- **Release cycles**: commit and issue activity often surges before a release
  and drops sharply after; this periodicity should be modelled rather than
  treated as noise.
- **Sprint velocity**: in agile projects, story points completed per sprint form
  a series with weekly seasonality and occasional spikes from unplanned work.
- **Post-release defect density curves**: defect reports rise in the first weeks
  after a release, then decay — a well-known empirical pattern.

## Changepoint Detection

A **changepoint** is a location in a time series where the statistical properties
(mean, variance, or trend) shift abruptly. In SE, changepoints often indicate:
a team restructuring, adoption of a new testing process, an architectural
overhaul, or a policy change.

The `changepoint` package in R provides PELT (Pruned Exact Linear Time) and
BinSeg algorithms for detecting single or multiple changepoints in mean or
variance.

```{r eval=FALSE}
library(changepoint)
# Detect changes in mean (e.g., number of weekly commits)
cpt_result <- cpt.mean(as.numeric(gmts), method = "PELT", penalty = "BIC")
plot(cpt_result)
cpts(cpt_result)  # positions of detected changepoints
```

For more gradual structural breaks, `strucchange` implements the Bai-Perron
framework for testing and estimating structural changes in linear regression
models.


## Anomaly Detection

Point anomalies (unexpected spikes or dips) in SE time series can indicate
data quality issues, incidents, or genuine development events. The `forecast`
package includes `tsoutliers()` for automatic outlier detection. The
`anomalize` package (part of the tidyverse-adjacent stack) provides a tidy
interface combining decomposition with anomaly scoring.

```{r eval=FALSE}
library(anomalize)
library(tibble)
library(dplyr)

# Example: detect anomalous commit weeks
gm_tbl <- tibble(
  date = seq(as.Date('2010-05-22'), as.Date('2015-11-16'), by = '1 week'),
  commits = as.numeric(gmts)
)

gm_tbl |>
  time_decompose(commits, method = "stl") |>
  anomalize(remainder, method = "iqr") |>
  plot_anomalies()
```


## Prophet for SE Forecasting

Facebook's **Prophet** model handles time series with strong multiple
seasonalities (weekly, annual), holiday effects, and missing data — all
common in SE repositories. It is particularly useful for projecting issue
volume, review load, or CI build queue length over a planning horizon.

```{r eval=FALSE}
# install.packages("prophet")
library(prophet)

df_prophet <- data.frame(
  ds = seq(as.Date('2010-05-22'), as.Date('2015-11-16'), by = '1 week'),
  y  = as.numeric(gmts)
)
m   <- prophet(df_prophet, weekly.seasonality = TRUE)
fut <- make_future_dataframe(m, periods = 52, freq = "week")
fct <- predict(m, fut)
plot(m, fct)
prophet_plot_components(m, fct)
```


## Further Resources on Time Series in R

- Hyndman & Athanasopoulos (2021) *Forecasting: Principles and Practice* (3rd ed.)
  — free online: <https://otexts.com/fpp3/>
- `tsibble`, `fable`, and `feasts` packages for tidy time series analysis
- CRAN Task View on Time Series:
  <https://cran.r-project.org/view=TimeSeries>
- A Little Book of R for Time Series:
  <https://a-little-book-of-r-for-time-series.readthedocs.io/>


