
# Social Network Analysis in SE


In this example, we will data from the MSR14 challenge. Further information and datasets:
[http://openscience.us/repo/msr/msr14.html](http://openscience.us/repo/msr/msr14.html)

Similar databases can be obtained using MetricsGrimoire or other tools.

In this simple example, we create a network form the users and following extracted from GitHub and stored in a MySQL database.

We can read a file directely from MySQL dump 

```{r  eval = FALSE}	
library(RMySQL)

# Connecting to MySQL
mydb = dbConnect(MySQL(), user='msr14', password='msr14', dbname='msr14', host='localhost')

# Retrieving data from MySQL
sql <- "select user_id, follower_id from followers limit 100;"
rs = dbSendQuery(mydb, sql)
data <- fetch(rs, n=-1)
```

Alternatively, we can create e CSV file directly from MySQL and load it

```{r eval = FALSE}

$mysql -u msr14 -pmsr14 msr14


> SELECT 'user','follower'
UNION ALL
SELECT user_id,follower_id 
	FROM followers 
	LIMIT 1000 
	INTO OUTFILE "/tmp/followers.csv"
	FIELDS TERMINATED BY ','
	LINES TERMINATED BY '\n';
```

```{r}
# Data already extracted and stored as CSV file (for demo purposes)
dat = read.csv("./datasets/sna/followers.csv", header = FALSE, sep = ",")
dat <- head(dat,100)
```

We can now create the graph

```{r}
library(igraph)
```

```{r}
# Create a graph
g <- graph.data.frame(dat, directed = TRUE)
```


Some values:
```{r}
summary(g); 
```


Plotting the graph:


```{r}
layout1 <-  layout.fruchterman.reingold(g)
plot(g, layout1)
```


Other layout
```{r}
plot(g, layout=layout.kamada.kawai)
```

A tk application can launched to show the plot interactively:
```{r eval = FALSE}
plot(g, layout = layout.fruchterman.reingold)
```


Some metrics:
```{r}
metrics <- data.frame(
  deg = degree(g),
  bet = betweenness(g),
  clo = closeness(g),
  eig = evcent(g)$vector,
  cor = graph.coreness(g)
)

#
head(metrics)
```



To fix and to do: Explain metrics and better graphs

```{r eval = FALSE}
library(ggplot2)

ggplot(
  metrics,
  aes(x=bet, y=eig,
    label=rownames(metrics),
    colour=res, size=abs(res))
)+
xlab("Betweenness Centrality")+
ylab("Eigenvector Centrality")+
geom_text()
+
theme(title="Key Actor Analysis")


V(g)$label.cex <- 2.2 * V(g)$degree / max(V(g)$degree)+ .2
V(g)$label.color <- rgb(0, 0, .2, .8)
V(g)$frame.color <- NA
egam <- (log(E(g)$weight)+.4) / max(log(E(g)$weight)+.4)
E(g)$color <- rgb(.5, .5, 0, egam)
E(g)$width <- egam
# plot the graph in layout1
plot(g, layout=layout1)
```

Further information:

http://sna.stanford.edu/lab.php?l=1


