
# Text Mining Software Engineering Data 

In software engineering, there is a lot of information in plain text such as requirements, bug reports, mails, reviews from applicatons,  etc.  Typically that information can be extracted from Software Configuration Management Systems (SCM), Bug Tracking Systems (BTS) such as Bugzilla or application stores such as Google Play or Apple's AppStore, etc. can be mined to extract relevant information. Here we briefly explain the text mining process and how this can be done with R.

The main package for _text mining_ is `tm` [@FeinererH15,@FeinererHM08]. Another popular package is `wordcloud`. 

```{r setup, message=FALSE, eval=FALSE, echo=FALSE, warning=FALSE}
# - Install packages

#pckgs_needed <- c("tm", "wordcloud")
#install.packages(pckgs_needed, dependencies = TRUE)
```

## Terminology 

We need to consider:

  * Importing data. A _Corpus_ is a collection of text documents, implemented as VCorpus (corpora are R object held in memory). The `tm` provides several corpus constructors: `DirSource`, `VectorSource`, or `DataframeSource` (`getSources()`). readerControl of the corpus constructor has to be a list with the named components reader and language
  
  * Preprocessing 
  
  * Inspecting and exploring data: Individual documents can be accessed via [[
  
  * Transformations: Transformations are done via the `tm_map()` function. 
    + `tm_map(_____, stripWhitespace)`  
    + `tm_map(_____, content_transformer(tolower))`
    + `tm_map(_____, removeWords, stopwords("english"))`
    + `tm_map(reuters, stemDocument)`
  * Creating `Term-Document` Matrices: TermDocumentMatrix and DocumentTermMatrix
     
  * Relationships between terms.
    + `findFreqTerms(_____, anumber)`
    + `findAssocs(Mydtm, "aterm", anumbercorrelation)`
    + A dictionary is a (multi-)set of strings. It is often used to denote relevant terms in text mining.  
    + `inspect(DocumentTermMatrix( newsreuters, list(dictionary = c("term1", "term2", "term3"))))`.   It displays detailed information on a corpus or a term-document matrix.

  * Clustering
    




## Example of classifying bugs from Bugzilla 


Bugzilla is Issue Tracking System that allow us to follow the evolution of a project.

The following example shows how to work with entries from Bugzilla. It is assumed that the data has been extracted and we have the records in a flat file (this can be done using Web crawlers or directly using the SQL database).


```{r readFileCompendiumARFF, message=FALSE, warning=FALSE}
library(foreign)
# path_name <- file.path("C:", "datasets", "textMining")
# path_name
# dir(path_name)

options(stringsAsFactors = FALSE)
d <- read.arff("./datasets/textMining/reviewsBugs.arff" )
str(d)
head(d,2)
# fifth entry
d$revContent[5]
d$revBug[5]
```

Creating a Document-Term Matrix (DTM)

```{r Create DTM, message=FALSE, warning=FALSE, include=FALSE}
library(tm)

ds <- DataframeSource(data.frame(doc_id=row.names(d),
                                 text=d$revContent))
dsc <- Corpus(ds)

# weighting=TfIdf weighting is Tf-Idf
# minWordLength=WL the minimum word length is WL
# minDocFreq=ND each word must appear at least in ND docs

# Other options of DTM
# These are not really needed, if preprocessing has been carried out:
# stemming=TRUE stemming is applied
# stopwords=TRUE stopwords are eliminated
# removeNumbers=TRUE numbers are eliminated

dtm <- DocumentTermMatrix(dsc, control = list(weighting = weightTfIdf, minDocFreq=3, stopwords = TRUE, removeNumbers = TRUE))

# dim(dtm)
# inspect(dtm) #[1:10,1:10])

# dtm.70=removeSparseTerms(dtm,sparse=0.7)
# dtm.70 # or dim(dtm.70)
# note that the term-document matrix needs to be transformed (casted)
# to a matrix form in the following barplot command

sparseparam <- 0.90
dtm_sprs <- removeSparseTerms(dtm,sparse=sparseparam)
maintitle <-paste0("Most frequent terms (sparseness=" ,sparseparam , "  )")
barplot(as.matrix(dtm_sprs),xlab="terms",ylab="number of occurrences", main=maintitle)
```


As data frame:

```{r message=FALSE,results='hide'}
#dtmdf <- as.data.frame(dtm.90)
#dtmdf <- as.data.frame(inspect(dtm_sprs))
dtmdf <- as.data.frame(as.matrix(dtm_sprs))
# rownames(dtm)<- 1:nrow(dtm)

class <- d$revBug
dtmdf <- cbind(dtmdf,class)
head(dtmdf, 3)
```

Now, we can explore things such as "which words are associated with "feature"?"

```{r}
# which words are associated with "bug"?
findAssocs(dtm, 'bug', .3)
```

And find frequent terms.

```{r frequentTerms, message=FALSE}
findFreqTerms(dtm,15)
```

Use any classifier now:

```{r tmCaret, message=FALSE}
library(caret)
library(randomForest)

inTraining <- createDataPartition(dtmdf$class, p = .75, list = FALSE)
training <- dtmdf[ inTraining,]
testing  <- dtmdf[-inTraining,]

fitControl <- trainControl(## 5-fold CV
                           method = "repeatedcv",
                           number = 5,
                           ## repeated ten times
                           repeats = 5)


gbmFit1 <- train(class ~ ., data = training,
                 method = "gbm",
                 trControl = fitControl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = FALSE)

gbmFit1

# trellis.par.set(caretTheme())
# plot(gbmFit1)
# 
# trellis.par.set(caretTheme())
# plot(gbmFit1, metric = "Kappa")

head(predict(gbmFit1, testing, type = "prob"))

conf_mat <- confusionMatrix(testing$class, predict(gbmFit1, testing))
conf_mat
```


And finally, a word cloud as an example that appears everywhere these days.

```{r WordCloud, message=FALSE}
library(wordcloud)

# calculate the frequency of words and sort in descending order.
wordFreqs=sort(colSums(as.matrix(dtm_sprs)),decreasing=TRUE)

wordcloud(words=names(wordFreqs),freq=wordFreqs)
```


## Extracting data from Twitter

The hardest bit is to link with Twitter. Using the TwitteR package is explained following this [example](./twitter.Rmd).


