
# Descriptive Statistics


## Normality 
  - A normal distribution is an arrangement of a data set in which most values cluster in the middle of the range
  - A graphical representation of a normal distribution is sometimes called a *bell curve* because of its shape.
  - Many procedures in statistics are based on this property. *Parametric* procedures require the normality property.
  - In a normal distribution about 95% of the probability lies within 2 Standard Deviations of the mean.
  - Two examples: one population with mean 60 and the standard deviation of 1, and the other with mean 60 and sd=4 (means shifted to 0)
  
```{r}
main.title <- "Area within 2 SD of the mean"
par(mfrow=c(1,2))
plot(function(x) dnorm(x, mean = 0, sd = 1),
xlim = c(-3, 3), main = "SD 1", xlab = "x",
ylab = "", cex = 2)
segments(-2, 0, -2, 0.4)
segments(2, 0, 2, 0.4)
plot(function(x) dnorm(x, mean = 0, sd = 4),
xlim = c(-12, 12), main = "SD 4", xlab = "x",
ylab = "", cex = 2)
segments(-8, 0, -8, 0.1)
segments(8, 0, 8, 0.1)
```

  - if we sample from this population we get "another population".
  
```{r}
sample.means <- rep(NA, 1000)
for (i in 1:1000) {
  sample.40 <- rnorm(40, mean = 60, sd = 4) #rnorm generates random numbers from normal distribution
  sample.means[i] <- mean(sample.40)
}
means40 <- mean(sample.means)
sd40 <- sd(sample.means)
means40
sd40
```
- These sample means are another "population". The sampling distribution of the sample mean is normally distributed meaning that the "mean of a representative sample provides an estimate of the unknown population mean".
```{r}
hist(sample.means)
```



## Getting the Data. Descriptive statistics.
  - Set the path to your files 
  - Read the Telecom1 dataset and print out the summary statistics with the command *summary*
```{r}
options(digits=3)
telecom1 <- read.table("../datasets/Telecom1.csv", sep=",",header=TRUE, stringsAsFactors=FALSE, dec = ".") #read data
summary(telecom1)
```
  - We see that this dataset has three variables (or parameters) and few data points (18)
    + size: the independent variable
    + effort: the dependent variable
    + EstTotal: the estimates coming from an estimation method
  - Basic Plots
```{r}
par(mfrow=c(1,2)) #two figures per row
size_telecom1 <- telecom1$size
effort_telecom1 <- telecom1$effort
hist(size_telecom1, col="blue")
lines(density(size_telecom1))
plot(density(size_telecom1))
hist(effort_telecom1, col="blue")
plot(density(effort_telecom1))

# violin plots for those two variables
library(vioplot)
vioplot(size_telecom1) 
vioplot(effort_telecom1)

boxplot(size_telecom1)
boxplot(effort_telecom1)

par(mfrow=c(1,1))
qqnorm(size_telecom1, main="Q-Q Plot of 'size'")
qqline(size_telecom1, col=2, lwd=2, lty=2) #draws a line through the first and third quartiles
qqnorm(effort_telecom1,  main="Q-Q Plot of 'effort'")
qqline(effort_telecom1)
```
  - We observe the non-normality of the data. 
  
  - We may look the posible relationship between size and effort with a scatterplot
  
```{r}
plot(size_telecom1, effort_telecom1)
```


## China dataset 


```{r}
library(foreign)
china <- read.arff("../datasets/china.arff")
china_size <- china$AFP
summary(china_size)
china_effort <- china$Effort
summary(china_effort)
par(mfrow=c(1,2))
hist(china_size, col="blue", xlab="Adjusted Function Points", main="Distribution of AFP")
hist(china_effort, col="blue",xlab="Effort", main="Distribution of Effort")
boxplot(china_size)
boxplot(china_effort)
qqnorm(china_size)
qqline(china_size)
qqnorm(china_effort)
qqline(china_effort)
```
  - We observe the non-normality of the data. 

## Normality. Galton data
  - It is the data based on the famous 1885 Francis Galton's study about the relationship between the heights of adult children and the heights of their parents.


```{r, echo=FALSE}
# library(UsingR); data(galton)
galton <- read.csv("../datasets/galton.csv")
par(mfrow=c(1,2))
hist(galton$child,col="blue") #,breaks=100)
hist(galton$parent,col="blue") #,breaks=100)
```


## Normalization

  - China dataset. Take logs in both independent variables. 
  
  
```{r, echo=FALSE}
par(mfrow=c(1,2))
logchina_size = log(china_size)
hist(logchina_size, col="blue", xlab="log Adjusted Function Points", main="Distribution of log AFP")
logchina_effort = log(china_effort)
hist(logchina_effort, col="blue",xlab="Effort", main="Distribution of log Effort")
qqnorm(logchina_size)
qqnorm(logchina_effort)
```


- If the log-log transformation is used the estimation equation is $$y= e^{b_0 + b_1 log(x)} $$

## China dataset. Correlation. 

  - **Correlation** is a statistical relationship between two sets of data.
  - With the whole dataset we may check for the linear Correlation of the variables we are interested in
  
  
```{r}
par(mfrow=c(1,1))
plot(china_size,china_effort)
cor(china_size,china_effort)
cor.test(china_size,china_effort)
cor(china_size,china_effort, method="spearman")
cor(china_size,china_effort, method="kendall")
```

