---
title: "Feature Selection Example"
output: html_document
---

Feature Selection



```{r}
library(caret)
library(doParallel) # parallel processing
library(dplyr) # Used by caret
library(pROC) # plot the ROC curve

library(foreign)

### Use the segmentationData from caret
# Load the data and construct indices to divided it into training and test data sets.
#set.seed(10)
kc1 <- read.arff("../datasets/defectPred/D1/KC1.arff")
```


```{r}
 inTrain <- createDataPartition(y = kc1$Defective,
  ## the outcome data are needed
  p = .75,
  ## The percentage of data in the
  ## training set
  list = FALSE)
```

The function `createDataPartition` does a stratified partitions.

```{r}
training <- kc1[inTrain,]
nrow(training)

testing <- kc1[-inTrain, ]
nrow(testing)
```


The train function can be used to
  + evaluate, using resampling, the effect of model tuning parameters on performance
  + choose the "optimal" model across these parameters
  + estimate model performance from a training set

```{r}
fitControl <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           ## repeated ten times
                           repeats = 10)


```
gbmFit1 <- train(Defective ~ ., data = training,
                 method = "gbm",
                 trControl = fitControl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = FALSE)
gbmFit1

```{r}
plsFit <- train(Defective ~ .,
 data = training,
 method = "pls",
 ## Center and scale the predictors for the training
 ## set and all future samples.
 preProc = c("center", "scale")
)

```

```{r}
testPred <- predict(gbmFit1, testing)
postResample(testPred, testing$Defective)
sensitivity(testPred, testing$Defective)
confusionMatrix(testPred, testing$Defective)
```

When there are three or more classes, confusionMatrix will show the confusion matrix and a set of "one-versus-all" results. 