---
title: "Script Assignment LR"
output: html_document
---


## Assignment. Create an R Markdown document with a basic Linear Regression method
The assignment consists of two parts
* A) Create the linear regression model between two variables: the predictor is "size" and the response is "effort"
* B) Optional: apply other models and compare different evaluation metrics

### Part A. Build a regression model for predicting "effort"

The assignment will consist of creating a R Markdown document that must run some of the R scripts given in class. Each student must select a dataset from the Promise repository or other source of software engineering data (the folder datasets in the R project). The student will have to analyze the dataset and extract some conclusions about the data and about the model proposed. The basic model required is the linear regression, but the student is free to add additional predictive models.

The steps that you have to follow in order to complete the assignment are:

1.  Create a New R project and a new .Rmd document

2.  Go to the "datasets/effortEstimation" folder and download one of the datasets related to effort, for example: boehm.arff, albrecht.arff, miyazaki.arff. The file should contain a variable "size" and other variable "effort"

3.  Read the dataset into the R environment. Adapt lines 88-92 of **document 200**. You can get the default path with the command `getwd() And you can build the string with the path to the file with the function paste0(string1,string2,...)
```{r dataset1, tidy=TRUE}
#assuming you are in the default directory for the project. Or copy and paste the file(s) in a folder in your project
path2thefile <- paste0(getwd(), "/datasets/effortEstimation/...thefile....csv")
path2thefile
thefile <- read.table(path2thefile, sep=",",header=TRUE, stringsAsFactors=FALSE, dec = ".") #read data
```

4.  Adapt lines 102-129 of **document 200** to the dataset that you've read

5.  Adapt lines 137-139: we plot the dataset
    Check for normality

6.  Transform the variables, if needed. Lines 180-190 in document 200.
    Use the function `log`
7.  Plot the transformed variables
8.  Check for correlation in the original variables. Lines 202-209 in document 200.
    We may check for correlation in the transformed variables, too.
9.  We start building the model now.
10. Split the dataset into Training and Testing . Lines 71-79 in document 440 show how it is done on the Telecom1 dataset. Repeat the same process with the file that you have read. 
11. Now you should have two files: one contains the training data and another the test data. We save those files as .csv files. Later we will build a data package for sharing them through a package stored in Github

```{r}
write.csv(thefile_train, "thefile_train.csv") #write data
write.csv(thefile_test, "thefile_test.csv") #write data
```
12. Build the linear model with the training data lines 81-91 in document 440. It is the same code used several times in document 417 (regression.Rmd). Example: lines 272-279 in document 417
13. Plot the linear model on the Training data. Same lines as above.
14. Evaluate the model on the Test data. If the variables have been transformed, your code should be similar to lines 95-107 in document 440
15. Compute classical measures MAR .... Lpred: lines 43-65 in document 440
16. Plot the model on the Test data and write some conclusions about the model.

* B) Optional: apply other models and compare different evaluation metrics

There are other models that could be easy added and compred.
Also, there are other metrics from statics that could be considered.
